---
title: "Homework 3"
author: Amanda Warnock
output: github_document
---



```{r}
library(tidyverse)
library(p8105.datasets)
data("instacart")

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r col(instacart)` columns. 

Observations are the level of items in order by user. There are user/order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

There are 134 aisles, and the most items are from fresh vegetables. 


Make a plot of only aisles with more than 10,000 items ordered. 

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle), 
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


Make a table for baking ingredients, dog food, and packaged vegetables/fruits.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

apples vs ice cream

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```




